---
title: "Respiration_calcs"
author: "Lydia Vaughn"
date: "May 4, 2018"
output: html_document
---

####1. Load packages
```{r}
library(tidyverse)
```

####2. Import all CO2 concentration files into a list of data frames
```{r}
path <- "data/collected_20181002" #change date as needed throughout code.  Also change the differently formatted date in chunk 7.

files <- list.files(path, full.names=TRUE) #list files in folder

#generate list of all CO2 concentration files
concentrations <- lapply(setNames(grep("CO2_", ignore.case=F, value=T, x=files), make.names(gsub("*.csv$", "", gsub("data.collected_20181002.", "", grep("CO2_", ignore.case=F, value=T, x=files))))), read.csv) #change date as needed

```

####3. Import the volume, key, and weights data frames
```{r}
list2env(lapply(setNames(grep("CO2_", ignore.case=F, value=T, x=files, invert=T), make.names(gsub("*.csv$", "", gsub("data/collected_20181002/", "", grep("CO2_", ignore.case=F, value=T, x=files, invert=T))))), read.csv), envir = .GlobalEnv)
```

####4. Import the sample inventory data frame
```{r}
inventory <- read.csv('data/sample_inventory.csv', header=T, sep=',', stringsAsFactors=F)
```

####5. convert the vol data frame to long format and format dates
```{r}
longvol <- gather(headspace_vol, key="description", value="vol_mL", 5:length(names(headspace_vol)))
longvol$date <- sapply(strsplit(as.character(longvol$description), "_"), "[", 2)
longvol$description <- sapply(strsplit(as.character(longvol$description), "_"), "[", 1)
```

####6. format dates (all data frames)
```{r}
longvol$date <- as.Date(longvol$date, "%Y%m%d")
inventory$date <- as.Date(as.character(inventory$collection_date), format = "%m/%d/%y")
concentrations <- concentrations %>% lapply(. %>% mutate(date = as.Date(date, "%m/%d/%y")))
```

####7. and add a variable for increment number to the key data frame
This isn't working well yet; it doesn't match the increments for the 20181002 data
```{r}
#key <- key %>% left_join(inventory %>% filter(date == "2018-03-31") %>% select("plot", "top_depth_cm", "bottom_depth_cm", "increment"))
#key <- key %>% left_join(inventory %>% filter(date == "2018-06-05") %>% select("plot", "top_depth_cm", "bottom_depth_cm", "increment"))
#key <- key %>% left_join(inventory %>% filter(date == "2018-08-07") %>% select("plot", "top_depth_cm", "bottom_depth_cm", "increment"))
key <- key %>% left_join(inventory %>% filter(date == "2018-10-01" | date == "2018-10-02") %>% select("plot", "top_depth_cm", "bottom_depth_cm", "increment"))
```

####8. calculate mass of C in the headspace on each measurement date
```{r}
mgC <- longvol %>% filter(description == "headspace") %>% select("ID", "vol_mL", "date") %>% left_join(bind_rows(concentrations %>% lapply(. %>% select("ID", "postflush", "calibrated_ppm", "date")))) %>% mutate(headspace_mgC = 4.98169e-7 * calibrated_ppm * vol_mL) %>% rename(headspace_mL = vol_mL)
```

####9. calculate the mass of C removed from the jar (in the IRGA sample) on each measurement date
```{r}
longvol$postflush <- ifelse(longvol$description == "removedpostflush", "yes", "no")
longvol$description <- ifelse(longvol$description == "removedpostflush", "removed", longvol$description)

mgC <- mgC %>% full_join(longvol %>% filter(description == "removed") %>% select("ID", "postflush", "vol_mL", "date") %>% left_join(bind_rows(concentrations %>% lapply(. %>% select("ID", "postflush", "calibrated_ppm", "date")))) %>% mutate(removed_mgC = 4.98169e-7 * calibrated_ppm * vol_mL) %>% rename(removed_mL = vol_mL))
```

####10/2 only: fill in an approximate CO2 concentration for 10/22, based on a linear increase between 10/19 and 10/25.  
Comment out this chunk for all sampling dates other than 10/2
```{r}
mgC <- longvol %>% filter(description == "headspace") %>% select("ID", "vol_mL", "date") %>% left_join(bind_rows(concentrations %>% lapply(. %>% select("ID", "postflush", "calibrated_ppm", "date")))) 

d1 <- data.frame(concentrations[3]) %>% filter(CO2_20181019.ID == "29")
d2 <- data.frame(concentrations[5]) %>% filter(CO2_20181025.ID == "29")

mgC$calibrated_ppm <- ifelse(mgC$ID == "29" & mgC$date == "2018-10-22", (d1[,4] + d2[,4]) / 2, mgC$calibrated_ppm)

mgC <- mgC %>% mutate(headspace_mgC = 4.98169e-7 * calibrated_ppm * vol_mL) %>% rename(headspace_mL = vol_mL)

mgC <- mgC %>% full_join(longvol %>% filter(description == "removed") %>% select("ID", "postflush", "vol_mL", "date")) %>% mutate(removed_mgC = 4.98169e-7 * calibrated_ppm * vol_mL) %>% rename(removed_mL = vol_mL)
```

####10. for each sample, calculate the additional C respired (since the previous measurement) and the cumulate C respired
```{r}
for(j in 1:length(unique(mgC$ID))){
  
  sub <- mgC %>% filter(ID == unique(mgC$ID)[j]) %>% arrange(date, postflush) 

  for(i in 1:length(sub$date)) {
    sub$accum_mgC[i] = sub$headspace_mgC[i] - sub$headspace_mgC[i-1] + sub$removed_mgC[i-1]
    }
  sub$accum_mgC <- ifelse(sub$date == min(sub$date), "0", sub$accum_mgC)
  if(j == "1") {accum <- sub}
  if(j != "1") {accum <- accum %>% full_join(sub)}
  }

accum <- accum %>% filter(!is.na(accum_mgC)) %>% group_by(ID) %>% mutate(cum_mgC = cumsum(accum_mgC)) %>% full_join(key)
```

####11. create a variable for number of days since the start of the incubation
```{r}
min(accum$date)
accum <- accum %>% group_by(ID) %>% mutate(days = date - min(date))
accum$days <- as.numeric(accum$days)
```

####12. calculate the respiration rate per g soil C (average over the whole 2-week incubation)
```{r}
rate <- accum %>% full_join(key) %>% full_join(weights %>% select(plot, top_depth_cm, wetsoil_g, gwc_drysoil)) %>% filter(days == 14, !is.na(wetsoil_g)) %>% mutate(drysoil_g = wetsoil_g * 100 / (gwc_drysoil + 100), rate_mgC_d_g = cum_mgC / days / drysoil_g) 

#range of CO2 production rate
min(rate$rate_mgC_d_g)
max(rate$rate_mgC_d_g)
```

####13. save the calculated respiration rates as a .csv file
```{r}
write.csv(rate %>% select("plot", "top_depth_cm", "bottom_depth_cm", "increment", "rate_mgC_d_g"), file = "data/calculated_respiration/respiration_rate_20181002.csv") 
```

####14. design plot theme
```{r}
plottheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text = element_text(color="black", size=12)) +
  theme(axis.title = element_text(size=14)) +
  theme(legend.position = c(.9, .85)) +
  theme(legend.title = element_text(size=14)) +
  theme(legend.text = element_text(size=12)) 
```

####15. plot cumulative respiration in each plot
```{r}
for(i in 1:length(unique(accum$ID))){
  sub <- accum %>% filter(ID == unique(accum$ID)[i])
  chamber <- paste(sub$plot[1], sub$top_depth_cm[1]) 
  chamber <- paste(chamber, sub$bottom_depth_cm[2], sep = "-") 
  plot <- ggplot(sub, aes(y = cum_mgC, x = days)) +
    geom_vline(xintercept = 14, lty = "dashed", color = "gray") +
    geom_point() +
    ylab("Cumulative respiration (mg C)") +
    xlab("Days") +
    ggtitle(chamber)
    #ggtitle(paste("sample", unique(accum$ID)[i]))
  print(plot + plottheme)
}
```

