---
title: "Temporal respiration plots"
author: "Lydia Vaughn"
date: "26/10/2018"
output: html_document
---

####1. Load packages
```{r}
library(tidyverse)
```

####2. Import and name all calculated respiration rate files 
```{r}
files <- list.files("data/calculated_respiration", full.names=TRUE) #list files in folder

list2env(lapply(setNames(files, make.names(gsub("*.csv$", "", gsub("data.calculated_respiration.", "", x = files)))), read.csv), envir = .GlobalEnv) #convert list of files to separate, named data frames
```

####3. Add a date identifier column to each data frame and combine them into one long data frame
```{r}
respiration_rate_20180331$date <- as.POSIXct("2018-03-31")
respiration_rate_20180605$date <- as.POSIXct("2018-06-05")
respiration_rate_20180807$date <- as.POSIXct("2018-08-07")
respiration_rate_20181002$date <- as.POSIXct("2018-10-02")

long <- respiration_rate_20180331 %>% full_join(respiration_rate_20180605) %>% full_join(respiration_rate_20180807) %>% full_join(respiration_rate_20181002)
```

####4. For each depth increment and date, summarize the respiration rate (mean, sd, n, and se)
```{r}
increments <- long %>% group_by(date, increment) %>% summarize(mean.resp = mean(rate_mgC_d_g)) %>% full_join(long %>% group_by(date, increment) %>% summarize(sd.resp = sd(rate_mgC_d_g))) %>% full_join(long %>% group_by(date, increment) %>% summarize(n.resp = length(rate_mgC_d_g))) %>% mutate(se.resp = sd.resp / n.resp^0.5)
```

####5. For each soil layer, plot the summarized data as a bar graph
```{r}
plottheme <- theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text = element_text(color="black", size=12)) +
  theme(axis.title = element_text(color="black", size=14)) +
  theme(legend.text = element_text(size=12)) +
  theme(legend.title = element_text(size=14)) +
  theme(plot.title = element_text(size = 14))

organic <- ggplot(increments %>% filter(increment == "1"), aes(y = mean.resp * 1000, x = as.factor(date))) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Organic soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
organic + plottheme

shallow <- ggplot(increments %>% filter(increment == "2"), aes(y = mean.resp * 1000, x = as.factor(date))) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Shallow mineral soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
shallow + plottheme

deep <- ggplot(increments %>% filter(increment == "3"), aes(y = mean.resp * 1000, x = as.factor(date))) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Deep mineral soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
deep + plottheme
```

####6. For each soil layer, plot mean respiration over time for each sample location
```{r}
plottheme2 <- theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text = element_text(color="black", size=12)) +
  theme(axis.title.y = element_text(color="black", size=14)) +
  theme(axis.title.x = element_blank()) +
  theme(legend.text = element_text(size=12)) +
  theme(legend.title = element_text(size=14)) +
  theme(plot.title = element_text(size = 14))


organic.byplot <- ggplot(long %>% filter(increment == "1"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Organic soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  #scale_x_continuous(labels = xlabels, breaks = c(1, 2, 3), limits = c(1, 3)) +
  scale_x_continuous(labels = c("3/31", "6/5", "8/7", "10/2"), breaks = c(1, 2, 3, 4), limits = c(1, 4))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(organic.byplot + plottheme2)

shallow.byplot <- ggplot(long %>% filter(increment == "2"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Shallow mineral soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  #scale_x_continuous(labels = xlabels, breaks = c(1, 2, 3), limits = c(1, 3)) +
  scale_x_continuous(labels = c("3/31", "6/5", "8/7", "10/2"), breaks = c(1, 2, 3, 4), limits = c(1, 4)) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(shallow.byplot + plottheme2)

deep.byplot <- ggplot(long %>% filter(increment == "3"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Deep mineral soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  #scale_x_continuous(labels = xlabels, breaks = c(1, 2, 3), limits = c(1, 3)) +
  scale_x_continuous(labels = c("3/31", "6/5", "8/7", "10/2"), breaks = c(1, 2, 3, 4), limits = c(1, 4)) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(deep.byplot + plottheme2)
```

####7. Remake the plots from the previous chunk, but without the 6/5 measurements, which I know were moisture-limited.  Instead of the previous formatting, label each line.
```{r}
organic.byplot.3 <- ggplot(long %>% filter(increment == "1" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Organic soil") +
  geom_point(aes(color = plot)) +
  geom_line(aes(color = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(organic.byplot.3 + plottheme2)

shallow.byplot.3 <- ggplot(long %>% filter(increment == "2" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Shallow mineral soil") +
  geom_point(aes(color = plot)) +
  geom_line(aes(color = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(shallow.byplot.3 + plottheme2)

deep.byplot.3 <- ggplot(long %>% filter(increment == "3" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Deep mineral soil") +
  geom_point(aes(color = plot)) +
  geom_line(aes(color = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(deep.byplot.3 + plottheme2)
```

####8. Make the above plots but this time don't label each line.
```{r}
organic.byplot.2 <- ggplot(long %>% filter(increment == "1" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Organic soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  #scale_x_continuous(labels = xlabels, breaks = c(1, 2, 3), limits = c(1, 3)) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(organic.byplot.2 + plottheme2)

shallow.byplot.2 <- ggplot(long %>% filter(increment == "2" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Shallow mineral soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(shallow.byplot.2 + plottheme2)

deep.byplot.2 <- ggplot(long %>% filter(increment == "3" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)), group = plot)) +
  ggtitle("Deep mineral soil") +
  geom_point(color = "#5272A3", alpha = 0.7) +
  geom_line(aes(group = plot), color = "#5272A3", alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(deep.byplot.2 + plottheme2)
```

####9. Remake the bar and line graphs, but separate data according to C/T
```{r}
long$CT <- sapply(strsplit(as.character(long$plot), ""), "[", 1)

increments.CT <- long %>% group_by(date, increment, CT) %>% summarize(mean.resp = mean(rate_mgC_d_g)) %>% full_join(long %>% group_by(date, increment) %>% summarize(sd.resp = sd(rate_mgC_d_g))) %>% full_join(long %>% group_by(date, increment) %>% summarize(n.resp = length(rate_mgC_d_g))) %>% mutate(se.resp = sd.resp / n.resp^0.5)

organic.CT <- ggplot(increments.CT %>% filter(increment == "1" & date != "2018-06-05"), aes(y = mean.resp * 1000, x = as.factor(date), fill = CT)) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Organic soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
organic.CT + plottheme

shallow.CT <- ggplot(increments.CT %>% filter(increment == "2" & date != "2018-06-05"), aes(y = mean.resp * 1000, x = as.factor(date), fill = CT)) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Shallow mineral soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
shallow.CT + plottheme

deep.CT <- ggplot(increments.CT %>% filter(increment == "3" & date != "2018-06-05"), aes(y = mean.resp * 1000, x = as.factor(date), fill = CT)) +
  geom_bar(stat="identity", position=position_dodge(0.8), width=0.75) +
  geom_hline(yintercept = 0, size = 0.25) +
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) +
  xlab("Sample collection date") +
  scale_x_discrete(breaks = c("2018-03-31", "2018-06-05", "2018-08-07", "2018-10-02"), labels = c("3/31", "6/5", "8/7", "10/2")) +
  ggtitle("Deep mineral soil") +
  geom_errorbar(aes(ymin = mean.resp * 1000 - se.resp * 1000, ymax = mean.resp * 1000 + se.resp * 1000), position = position_dodge(0.8), width = 0.2, size = 0.4) 
deep.CT + plottheme

organic.byCT.line <- ggplot(long %>% filter(increment == "1" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Organic soil") +
  geom_point(aes(color = CT)) +
  geom_line(aes(color = CT, group = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(organic.byCT.line + plottheme2)

shallow.byCT.line <- ggplot(long %>% filter(increment == "2" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Shallow mineral soil") +
  geom_point(aes(color = CT)) +
  geom_line(aes(color = CT, group = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(shallow.byCT.line + plottheme2)

deep.byCT.line <- ggplot(long %>% filter(increment == "3" & date != "2018-06-05"), aes(y = rate_mgC_d_g * 1000, x = as.numeric(as.factor(date)))) +
  ggtitle("Deep mineral soil") +
  geom_point(aes(color = CT)) +
  geom_line(aes(color = CT, group = plot), alpha = 0.4) +
  scale_x_continuous(labels = c("3/31", "8/7", "10/2"), breaks = c(1, 2, 3), limits = c(1, 3))+
  ylab(expression(atop(CO[2]~production, (mu*g~C~~g~soil^-1~~d^-1)))) 

print(deep.byCT.line + plottheme2)
```