---
title: "Water content"
author: "Lydia Vaughn"
date: "June 14, 2018"
output: html_document
---

```{r}
samples <- read.csv('data/sample_inventory.csv', header=T, sep=',', stringsAsFactors=F)
water_content <- read.csv('data/water_content.csv', header=T, sep=',', stringsAsFactors=F)
capacity <- read.csv('data/water_capacity_v2.csv', header=T, sep=',', stringsAsFactors=F)
```

```{r}
library(tidyverse)
```

####Add a variable for core increment to both the water content and capacity data frames
```{r}
capacity <- capacity %>% left_join(samples %>% select(collection_date, plot, top_depth_cm, bottom_depth_cm, increment)) 

water_content <- water_content %>% left_join(samples %>% select(collection_date, plot, top_depth_cm, bottom_depth_cm, increment)) 
```

####Make a dataframe that includes just sample info, gwc, and water-holding capacity
####Use it to calculate the percentage of maximum water holding capacity and the water needed to reach 65% of water holding capacity (in mL of water per g of soil)
```{r}
combined <- capacity %>% select(plot, increment, water_holding_capacity_percent_dryweight) %>% full_join(water_content %>% select(collection_date, plot, top_depth_cm, bottom_depth_cm, increment, gwc_drysoil)) %>% mutate(percent_whc = gwc_drysoil / water_holding_capacity_percent_dryweight * 100, add_water_mL_g = (0.65 * water_holding_capacity_percent_dryweight - gwc_drysoil) / 100)
```

####Import dataframe of starting (field moisture) sample weights used in the incubation
```{r}
#change file and path as needed
masses <- read.csv('data/collected_20180605/starting_weights_20180605.csv', header=T, sep=',', stringsAsFactors=F)
```

####Use GWC and additional water info to calculate sample-specific water additions
```{r}
water <- masses %>% left_join(combined %>% select(plot, top_depth_cm, bottom_depth_cm, collection_date, gwc_drysoil, percent_whc, add_water_mL_g)) %>% mutate(drysoil_g = wetsoil_g / (gwc_drysoil / 100 + 1), add_water_mL = add_water_mL_g * drysoil_g) %>% select(ID, drysoil_g, gwc_drysoil, percent_whc, add_water_mL)
```

####write dataframe
```{r}
write.csv(water, file = "data/collected_20180605/water_additions_20180605.csv") 

```

